Notes and Problems from My Research\vspace{-1ex}

### 2019-03-15 Hopf Algebra Notes/Hopf Algebras.tex
Update work

\subsection{Noncommutative Reflections}
\begin{defn}
	Let $g\in\Aut_{gralg}(A)$. Then
	\[\operatorname{Tr}(g)=\sum_0^\infty \operatorname{tr}(g|_{A_i})t^i\in k[[t]].\]
\end{defn}
Then a result (which?) gets us that $\operatorname{Tr}(g)=\frac{1}{p(t)}$ for a polynomial $p$.
Consider $\tilde A=k_{p_{ij}}[x_1,\dots,x_n]$ where $x_jx_i=p_{ij}x_ix_j$ for all $i<j$. Then 
\begin{defn}
	$g\in\Aut_{gralg}(\tilde A)$ is a reflection if $\operatorname{Tr}(g)=\frac{1}{(1-t)^{n-1}(1-\lambda t)}$.
\end{defn}
\end{document}

>>>>

\subsection{Noncommutative Reflections}
\begin{defn}
	Let $g\in\Aut_{gralg}(A)$. Then
	\[\operatorname{Tr}(g)=\sum_0^\infty \operatorname{tr}(g|_{A_i})t^i\in k[[t]].\]
\end{defn}
Then a result (which?) gets us that $\operatorname{Tr}(g)=\frac{1}{p(t)}$ for a polynomial $p$.
Consider $\tilde A=k_{p_{ij}}[x_1,\dots,x_n]$ where $x_jx_i=p_{ij}x_ix_j$ for all $i<j$. Then 
\begin{defn}
	$g\in\Aut_{gralg}(\tilde A)$ is a reflection if $\operatorname{Tr}(g)=\frac{1}{(1-t)^{n-1}(1-\lambda t)}$.
\end{defn}

\section{March 13, 2019}
Today we are going to discuss why considering algebras over $A$ is not quite sufficient in many cases --
why we may be instead interested in studying $_A\calM$.

Let $A$ be an algebra and $V$ a Hopf algebra. Let $V$ act on $A$ and define
\[A^V=\{a\in A|va=\varepsilon(v)a,\forall v\in V\}\]
We can see that we get $A^v=\Hom_V({_Vk},A)$ and then use this to compute the cohomology
\[\Ext_V^i(k,A)\]

We forgot the
\subsection{(Non)-Examples of the Day}
\begin{ex}
	Let $A$ be an algebra, $n\ge 2$ an integer. Then $M_n(A)$ can't be a Hopf algebra.

	The idea here is that there is no counit $\varepsilon:M_n(A)\to k$.
\end{ex}
\begin{ex}
	If $V_1$ and $V_2$ are Hopf algebras, then $V_1\oplus V_2$ is not a Hopf algebra.
	Notice here that it is both an algebra and a coalgebra!

	Here, use that $1_V=1_{V_1}+1_{V_2}$ and compute $\Delta(1_V)\ne 1_V\otimes 1_V$.
\end{ex}
\begin{ex}
	If $A$ is finite dimensional and not Frobenius, then $A$ can't be a Hopf algebra.
\end{ex}

\subsection{Some Category Theory}
Let $k$ be af field.
\begin{defn}
	Let $\mathscr C$ be a category. $\mathscr C$ is called \textbf{$k$-linear} if
	\begin{itemize}
		\item $\Hom_{\mathscr C}(M,N)\in\Vectk$ for all $M$ and $N$ in $\mathscr C$
		\item Composition is $k$-linear.
	\end{itemize}
\end{defn}
\begin{defn}
	A category $\mathscr C$ is called \textbf{finite} if $\mathscr C\cong {_A\calM_{fd}}$ 
	for some finite dimensional algebra $A$.
\end{defn}
\begin{defn}
	A category $\mathscr C$ is called \textbf{monoidal} (notice that sometimes this books 
	call this a tensor category) if it is equipped with
	\begin{itemize}
		\item A bifunctor $-\otimes -:\mathscr C\times\mathscr C\to \mathscr C$
		\item An associativity isomorphism: for all $X,Y,X\in\mathscr C$:
		\[a_{X,Y,Z}:(X\otimes Y)\otimes Z\xrightarrow{\sim} X\otimes(Y\otimes Z)\]
		\item An identity object $\mathbf{1}\in\mathscr{C}$
		\item An isomorphism $i:\mathbf{1}\otimes\mathbf{1}\to \mathbf{1}.$
	\end{itemize}
	That satisfies the pentagon axiom (Figure~\ref{fig:pentagon}) and the unit axioms
	\[L_\mathbf{1}X\xrightarrow{\sim} \mathbf{1}\otimes X,\qquad R_\mathbf{1}:X\xrightarrow{\sim}X\otimes \mathbf{1}\]
\end{defn}
\begin{figure}\label{fig:pentagon}
	\centering
	\begin{tikzcd}
		((W\otimes X)\otimes Y)\otimes Z\ar[d,"a_{W,X,Y}\otimes \id_Z"]\ar[r,"a_{W\otimes X,Y,Z}"] & (W\otimes X)\otimes (Y\otimes Z)\ar[dd,"a_{W,X,Y\otimes Z}"]\\
		(W\otimes(X\otimes Y))\otimes Z \ar[d,"a_{W,X\otimes Y,Z}"]& \\
		W\otimes((X\otimes Y)\otimes Z)\ar[r,"\id_W\otimes a_{X,Y,Z}"] & W\otimes(X\otimes (Y\otimes Z))
	\end{tikzcd}
	\caption{The Pentagon Axioms}
\end{figure}
\begin{ex}
	Let $A$ be an algebra and let $\mathscr C= ({_A\calM_A},\otimes_A,\mathbf{1}=A)$.
	Clearly this works. :)
\end{ex}

Recall that if $V$ is a Hopf algebra and $M\in {_V\calM_{fd}}$, then $M^*:=\Hom_k(M,k)\in\calM_V$
then we defined $M^{*S}=\tilde S\circ M^*\in {_V\calM}$. This is the right dual.

We can also define the left dual if $S$ is bijective. In this case, we define $M^{*S^{-1}}=\tilde S^{-1}\circ M^*\in{_V\calM}$.

Now let $(\mathscr C,\otimes, \mathbf{1})$ be a monoidal category.
\begin{defn}
	Let $X\in\mathscr C$. An object $X^*$ in $\mathscr C$ is said to be a \textbf{left dual of $X$}
	if there exist an \textbf{evaluation map}
	\[\operatorname{ev}_X:X^*\otimes X\to \mathbf{1}\]
	and a \textbf{coevaluation map}
	\[\operatorname{coev}_X:\mathbf{1}\to X\otimes X^*\]
	such that
	\[X\xrightarrow{\operatorname{coev}_X\otimes\id_X}(X\otimes X^*)\otimes X\to X\otimes (X^*\otimes X)\xrightarrow{\id_X\otimes \operatorname{ev}_X} X\]
	is an isomorphism as well as the map ``on the other side''.
\end{defn}
\begin{defn}
	In the context of the above, a \textbf{right dual of $X\in\mathscr C$} is a left dual of $X$ in $\mathscr C^{op}$.
	The right dual is denoted (confusingly) by ${^*X}$.
\end{defn}
So in the Hopf algebra case, ${^*M}=M^{*S}$ and $M^*=M^{*S^{-1}}.$
\begin{defn}
	A category $\mathscr C$ is called \textbf{rigid} if every $X\in\mathscr C$ has both left and right duals.
\end{defn}
\begin{defn}
	$\mathscr C$ is called a \textbf{tensor category} if 
	\begin{itemize}
		\item $\mathscr C$ is finite (this implies $k$-linear)
		\item It is monoidal
		\item it is rigid
		\item $\End_\mathscr{C}(\mathbf{1})=k$
	\end{itemize}
\end{defn}
\begin{rmk}
	If we remove the last condition and replace it with $\End(\mathbf{1})=\oplus k$, then
	we call $\mathscr C$ \textbf{multi-tensor.}
\end{rmk}
\begin{defn}
	$\mathscr C$ is called \textbf{fusion} if 
	\begin{itemize}
		\item It is tensor
		\item It is semisimple (every object is a direct sum of simple objects).
	\end{itemize}
\end{defn}
\begin{rmk}
	If instead $\mathscr C$ is multitensor and semisimple, then it is called \textbf{multifusion.}
\end{rmk}
\begin{ex}
	\begin{enumerate}
		\item $\Vectk$ is an example.
		\item If $V$ is a finite dimensional Hopf algebra, ${_V\calM}$ is tensor.
		\item If $V$ is also semisimple, ${_V\calM}$ is fusion.
	\end{enumerate}
\end{ex}
Recall this conjecture for the first day (Etinghof-Ostrik restated):
\begin{conj}
	If $\mathscr C$ is a tensor category then $\Ext^*(\mathbf{1},\mathbf{1})$ is Noetherian.
\end{conj}

\section{March 15, 2019}
On Wednesday, we said that if $M\in{_V\calM_{fd}}$ that $M^*=M^{*S^{-1}}$ and ${^*M}=M^{*S}$,
but in fact these are reversed.
\subsection{Example of the Day}
\begin{ex}
	Usually we have been defining Hopf algebras over vector spaces, but this will be an object in $\text{\sc grVect}_k$,
	that is $V=\oplus_{i\in\Z}V_i$ and if $M=\oplus M_i$ and $N=\oplus N_i$, then $M\otimes N=\oplus_i(\oplus_{s+t=i}M_s\otimes N_t)$.

	Now putting the ``symmetric'' back into the monoidal category, we define $\tau:M\otimes N\to N\otimes M$ via
	\[m\otimes n\mapsto (-1)^{|m||n|}n\otimes m.\]
	Then we have our symmetric monoidal category and we can define a Hopf algebra in it.

	$V=\Lambda(x_1,\dots,x_n)$, the exterior algebra on $n$ elements. Define the coalgebra structure 
	so that each $x_i$ is primitive. Note that here $V$ is a Hopf algebra over graded vector spaces, but 
	NOT over vector spaces (unless $\ch k=2$).

	To see this, we can compute 
	\[0=\Delta(x_i^2)=\Delta(x_i)^2=(x_i\otimes 1+1\otimes x)^2=x_i^2\otimes 1+x_i\otimes x_i+\tau(x_i\otimes x_i)+1\otimes x_i^2\]
	and if $\tau$ is the usual vector space twist (and the characteristic isn't 2) we get a contradiction.
\end{ex}

\subsection{Some History}
Before 1941 people studied rings like $kG, \mathcal U(\g)$ and $\mathcal O(G)$. An open question at that
point was ``If $\mathcal U(\g)$ is Noetherian, then is $\dim \g<\infty?$''

In 1941, Hopf algebras were born, although they were originally defined (by Hopf) to be over graded
vector spaces. Eventually people started considering them over vector spaces.

In 1967, Sweedler wrote the first book about Hopf algebra over vector spaces. Then later, near 1970,
Rota began working more on combinatorial Hopf algebras. It ends up in this context that often the 
M\"obius function ends up being the co(unit? multiplication?).

In 1975, Kaplansky posited 10 conjectures. Of those only one more remains:
\begin{conj}{Kaplansky, `75 (\#6)}
	If $V$ is a semisimple Hopf algebra and $M$ is a simple $V$-module, then 
	\[\dim_kM|\dim_kV.\]
\end{conj}

From 1960-90, the studies of $kG, \mathcal U(G)$ and $\mathcal O(G)$ and semisimple Hopf algebra continued,
led in large part by Radford. In 1987, Drinfeld and Jimbo came out with quantum groups.
At the same time Drinfeld introduced quasi- and braided Hopf algebras and weak Hopf algebras.

Of course we all know about Nichols algebras. Also we switched to different tensor categories, including
fusion, pivotal, and spherical ones.

In 1993 Lusztig invented the small quantum groups, which are of particular interest to representation theorists.
`97 was Brown-Goodearl. Then around 1990 there were lots of classification projects including
the one by Anruskiewitsch-Schneider project. Near this time Zhu proved that every Hopf algebra of dimension $p$
is a group algebra $k\Z/p$.

In 2004 was Etingof-Ostrik. The next year, Etingof-Nikshych-Ostrik wrote a big paper ``On Fusion Categories''
which really shifted the perspective from Hopf algebras to things in fusion categories, where many of the big
results were transferred.

In 2015 Etingof, Nikshych, Ostrik (and someone else) wrote ``Tensor Categories''.

\subsection{Extensions}
We can look at extensins of Brown-Goodearl. The question ``Is every Noetherian X Gorenstein?'' can be extended to be asked about any noun 
in \{Hopf Algebras, weak Hopf algebras, braided hopf algebras, Nichols algebgras, quasi-Hopf algebras\}.

We can also consider extensions of Etingof-Ostrik: ``Let $V$ be a finite dimensional X and $A$ be a 
Notherian left $V$-module such that $A^V$ is Noetherian. Is then $\Ext_V^*(k,A)$ Noetherian?'' Again
we can take $X$ to be anything in the above set.

In 2019, who knows!

\end{document}

### 2019-03-15 Lie Algebra Notes/Lie Algebras.tex
Update work

%characteristic
%\DeclareMathOperator{\char}{char}
\DeclareMathOperator{\ch}{ch}

@@

%Lie algebra stuff
\DeclareMathOperator{\gl}{\mathfrak{gl}}
\let\sl\relax
\DeclareMathOperator{\sl}{\mathfrak{sl}}
\DeclareMathOperator{\so}{\mathfrak{so}}
\let\sp\relax
\DeclareMathOperator{\sp}{\mathfrak{sp}}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\ad}{ad}
\DeclareMathOperator{\Der}{Der}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Rad}{Rad}
\let\t\relax
\DeclareMathOperator{\t}{\mathfrak{t}}
\DeclareMathOperator{\n}{\mathfrak{n}}
\DeclareMathOperator{\h}{\mathfrak{h}}
\DeclareMathOperator{\g}{\mathfrak{g}}
\DeclareMathOperator{\diag}{diag}

@@

%TODO: FIX CHAR/CH
\end{document}

>>>>

%characteristic
%\let\char\relax
%\DeclareMathOperator{\char}{char}
\DeclareMathOperator{\ch}{ch}

@@

%Lie algebra stuff
\DeclareMathOperator{\gl}{\mathfrak{gl}}
\let\sl\relax
\DeclareMathOperator{\sl}{\mathfrak{sl}}
\DeclareMathOperator{\so}{\mathfrak{so}}
\DeclareMathOperator{\su}{\mathfrak{su}}
\let\sp\relax
\DeclareMathOperator{\sp}{\mathfrak{sp}}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\ad}{ad}
\DeclareMathOperator{\Der}{Der}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Rad}{Rad}
\let\t\relax
\DeclareMathOperator{\t}{\mathfrak{t}}
\DeclareMathOperator{\n}{\mathfrak{n}}
\DeclareMathOperator{\h}{\mathfrak{h}}
\DeclareMathOperator{\g}{\mathfrak{g}}
\DeclareMathOperator{\diag}{diag}

@@

\section{March 13, 2019}
One correction from last week: the compact from of $\sp_{2n}(\C)$ is given by the skew-adjoing matrices of trace zero relative to the \textit{Hermitian}
form $(v,w)=\sum_iv_i\bar w_i$ on $\mathbb{H}^n$ (not skew-Hermitian).

\subsection{Continuing with Real Forms}
Now we will address the non-compact forms. We've seen the compact and split real forms, which form 
the opposite ends of a spectrum and now we consider the ones that lie in the middle.

\subsubsection{Real forms of type $A$}
In type $A$, we have already
seen $\su_n(\C)$ consisting of skew-adjoint matrices of trace zero relative to the positive definite Hermitian form on $\C$.
We've also seen $\sl_n(\R)$.

In between, we have $\su(p,q)$, consisting of the skey adjoint matrixes with respect to the Hermitian form $(-,-)_{pq}$ where
\[(v,w)=\sum_1^pv_i\bar w_i-\sum_{p+1}^{p_q}v_i\bar w_i\]
which is Hermitian of \textbf{signature} $(p,q)$ on $\C^{p+q}=\C^n$. WHen $p=n$ and $q=0$, we get the compact form $\su_n(\C)$.

We get one more real form -- only in even dimensions -- $\su^*(2n)$: all $n\times n$ matrices over the quaternions $\mathbb{H}$ whose trace is a real
combination of $i,j,k$.

\subsubsection{Types $B$ and $D$}
Here the story is mostly the same for these types with a small difference: we have $\so(p,q)$
in either, which is all skew-adjoint matrices with respect of the symmetric form $(-,-)_{pq}$
on $\R^{p+q}$ where
\[(v,w)=\sum_1^pv_iw_i-\sum_{p+1}^{p+q}v_iw_i.\]

But we get one extra in type $D$: $\so^*(2n)$. This consits of all skew-adjoint matrices with respect to the skey Hermitian form
\[(v,w)=\sum_1^n v_ij\bar w_i\]
where $j\in\mathbb{H}$.

\subsubsection{Type $C$}
Here we get $\sp(p,q)$ -- the skew-adjoint matrices with respect to the form as usual ($v_i\bar w_i$).
We also have $\sp(n,\R)$, the compact form.

Finally we get $\sp(2n,\R)$, the real matrices skew adjoint with respect to a skew-sypmmetric form on $\R^{2n}$.

\subsection{Cartan Involution}
In general, recall from last time that any real semisimple Lie algebra admits a Cartan decomposition
$L=K+P$ where $K$ is compact and $[P,P]\subseteq K$. Furthermore, the Killing form $\kappa$ restricts
to $K$ as a negative definite form and to $P$ as a positive definite form. Finally, $P$ is a $K$-sumbodule:
$[KP]\subseteq P$.

It follows that we get an involution on $L$ by decreeing that $\theta=1$ on $K$ and $\theta=-1$ on $P$ (???) the \textbf{Cartan
involution.} Moreover $L$ admits a complexification $L\otimes_\R\C=L'$, which is complex semisimple, and
$\theta$ extends to an involution of $L'$. We classify all real forms up to isomorphism by classifying the involutions
$\theta$ up to conjugacy. It ends up there are only finitely many real forms.

\subsection{Exceptional Cases}
It turns out htat in the exceptional cases we can label real forms explicitly as $X_N(a)$ where the real form of type $X_N$ has $a=\dim P-\dim K$.
Here the compact form is always $X_N(-\dim_CX_N)$
\subsubsection{$E_6$}
This ends up being the most interesting type (although it sits inside of its bigger brother).
Here we get the real forms:
\begin{figure}[h!]
	\centering
	\begin{tabular}{c|c|c|c}
		$E_6(6)$ & $E_6(2)$ & $E_6(-14)$ & $E_6(-26)$\\\hline
		$C_4$ & $A_5\times A_1$ & $D_5\times \C$ & $F_4$
	\end{tabular}
\end{figure}
where here the $\C$ refers to the fact that there will be a (one-dimensional) nontrivial center.
\subsubsection{$E_7$}
\begin{figure}[h]
	\centering
	\begin{tabular}{c|c|c}
		$E_7(7)$ & $E_7(-5)$ & $E_7(-25)$ \\\hline
		$A_7$ & $D_6\times A_1$ & $E_6\times \C$
	\end{tabular}
\end{figure}
\subsubsection{$E_8$}
\begin{figure}[h]
	\centering
	\begin{tabular}{c|c}
		$E_8(8)$ & $E_8(-24)$ \\\hline
		$D_8$ & $E_7\times A_1$
	\end{tabular}
\end{figure}
\subsubsection{$F_4$}
\begin{figure}[h!]
	\centering
	\begin{tabular}{c|c}
		$F_4(4)$ & $F_4(-20)$ \\\hline
		$C_3\times A_1$ & $B_4$
	\end{tabular}
\end{figure}
So again we get two.
\subsubsection{$G_2$}
Here we only have one form: $G_2(2)=A_1\times A_1$.

\subsection{Diagram Automorphisms}
There are three types of Lie algebras where diagram automorphisms of order two occur: namely 
$A_n$, $D_n$ and $E_6$. In these cases, we create a different kind of extended Dynkin diagram:
we add a vertex representing the lowest weight of the quotient representation that is formed
by beginning by looking at $L$ and then quotienting out by the fixed points of the diagram 
automorphism.

\subsection{Next time...}
The last lecture of the quarter will be dedicated to what happens when $k$ has positive characteristic.
There is a classification, but it is much longer and mroe complicated than the one we have given.

In particular, one might at first think that at least the split form of a complex semisimple Lie algebra
$L$ would have a characteristic $p$ analog. While this is true, it is not always (even) semisimple. 
The problem is that we completely lose the notion of positivity in positive characteristic, so 
the Killing form, even when restricted to a Cartan subalgebra, needn't be nondegenerate.

In the simplest case, consider $\ch k=p$ and the Lie algebra $\sl_n(k)$. The identity matrix
then has trace zero! Problems abound. Finally we will talk about the finite simple groups of Lie type,
which form the building blocks for the Lie groups we are interested in.

\section{March 15, 2019}
Let $L$ be a complex simple Lie algebra. We have seen that $L$ has a Chevalley basis
consisting of a basis of $H$, a Cartan subalgebra, with root system $\Phi$, and for each
(nonzero) root $\alpha\in\Phi$ a vector $x_\alpha\in L_\alpha$ such that whenever $\alpha,\beta,\alpha+\beta$
are all roots, $[x_\alpha,x_\beta]=c_{\alpha\beta}x_{\alpha+\beta}$. Then $c_{\alpha\beta}=-c_{-\alpha.-\beta}$
and $c_{\alpha\beta}=\pm(r+1)$ where $r$ is the largest integer (possibly zero) such that $\beta-r\alpha\in\Phi$.

Then the span $L_\Z$ over the integers of a Chevalley basis is a Lie algebra ``over $\Z$'', which is 
simple in the sense that any nonzero ideal comtains $nL_\Z$ for some nonzero $n\in\Z$. We have seen however that
the reduction $L_p$ of $L_\Z$ modulo $p$ need not be simple.

So we focus on the \textbf{group} generated by $\exp\ad cx_\alpha$ for $c\in\Z$ and $\alpha\in\Phi$.
Suppose that $\alpha,\beta,\beta+\alpha,\beta+2\alpha,\dots,$ are all roots. Then
\[\exp\ad x_\alpha(x_\beta)=\pm(r+1)(r+2)\cdots x_{\beta+n\alpha}\]
(since all root strings are actually finite, the series actually can be considered to be
finite in any one case). Thus $\pm(r+1)\cdots(r+n)/w!$ is an integer, so the terms in the series
expainsion of $\exp\ad cx_\alpha$ all preserve the $\Z$-span $L_\Z$ whence there reductions modulo
a prime $p$ preserve the corresponding span $L_p$ so that if $F$ is any field containing $\Z_p$
and if we set $L_F=L_p\otimes_{\Z/p}F$ then the group $G_F$ generated by $\exp\ad cx_\alpha$ acts on $L_F$.

Specializing down to the case when $F=\mathbb{F}_q$, the finite field of order $q=p^n$,
we therefore get a finite group acting on $L_q:=L_{\mathbb{F}_q}$. Then Chevalley proved that $L_q$ is always simple as a group, 
with just a few exceptions occurring for very small fields $F$, thereby exhibiting a number 
of families of finite simple groups, now said to be \textbf{of Lie Type.} Many of the resulting simple groups were known
before but Chevalley was the first to give a unified treatment.

\subsection{Examples}
An explicit example comes from $GL_n(\mathbb{F}_q)$. It can be shown (see many old prelims) to 
have order $q^{\binom{n}{2}}(q^n-1)\cdots(q-1)$. This is actually never simple! But luckily we want some smaller
groups! We have $SL_n(\mathbb{F}_q)$ which has order $q^{\binom{n}{2}}(q^n-1)\cdots(q^2-1)$.

Still not simple, unfortunately, but if we mod out scalars to $PSL_n(\mathbb{F}_q)$, we get a simple
group (whenever $q\ge 4$) of order $q^{\binom{n}{2}}(q^n-1)\cdots(q^2-1)/g$ where $g=\operatorname{gcd}(n,q-1)$.

We notice that the formula for the order of the simple group $PSL_n(\mathbb{F}_q)$ is given 
by a power of $q$ times a product of $q^i-1$ for varius $i$, divided the by the order of a central
subgroups.

Chevalley showed this pattern always holds: the exponent of $q$ is the dimension of the Borel
subalgebra of the Lie algebra giving rise to the group. The $q^i-1$ powers are called the \textbf{exponents}
of the Weyl group of the Lie algebra.

\subsection{Classificiations}
Let us look at all finite simple roups of nonprime order: we have
\begin{itemize}
	\item $A_n$, the alternating groups, for $n\ge 5$
	\item Finite groups of Lie type 
	\item Chevalley groups
	\item Further anaglogues of these arising from a nontrivial diagram automorphisms fo the Lie algebra
	\item Steinberg analogues on characteristic 2 and 3 on types $B_2$ and $G_2$
	\item Ree, Suzuki, and Tits (are these the discoverers of above?)
	\item Only finitely many others! 26 sporadic groups with the monster group being the largest.
\end{itemize}
\begin{rmk}
	There is also a monster Lie algebra.
\end{rmk}

Moreover, one can form an analgof of $L_\Z$ for the enveloping algebra: it has a $\Z$-form
$U_\Z$ consisting of all sums of products $\frac{x^{n_\alpha}_{-\alpha}}{n_\alpha!}\cdots\frac{x_\alpha^{n_\alpha}}{n_\alpha!}$
and this is stable under the action of the Chevalley group.

I missed the last bit because Monty was talking, but that wraps it up for the quarter!
\end{document}

### 2019-03-15 Reading Notes/reading_notes.tex
Update work

@@

@@

\section{To be filed:}
The last paper on Julia's page is a preprint with her and Dave Benson. There is a spectral sequence computation that might be illustrative.
There is also an older paper...

>>>>

\subsection{Week 8}
The last paper on Julia's page is a preprint with her and Dave Benson. There is a spectral sequence computation that might be illustrative.
There is also an older paper...

\subsection{Week 10}

@@

One of the first propositions (3.1) of the paper computes the cohomology of $\bbG_a^-$ with 
coefficients in $k$ (actually, refers to another paper) and computes how the ($\bbZ/2$-graded)
Steenrod operations act on it.

\begin{defn}[Steenrod Operations]
	If $p$ is odd and $A$ is a $\bbZ$-graded cocommutative Hopf algebra, then there are natural maps/operations
	\[\mathcal P^i:H^{s,t}(A,k)\to H^{s+(2i-t)(p-1),pt}(A,k)\]
	\[\beta \mathcal P^i:H^{s,t}(A,k)\to H^{s+1+(2i-t)(p-1),pt}(A,k)\]
	satisfying the properties
	\begin{enumerate}
		\item $\mathcal P^i=0$ whenever $2i<t$ or $2i>s+t$,
		\item $\beta\mathcal P^i=0$ whenever $2i<t$ or $2i\ge s+t$.
		\item $\mathcal P^i(x)=x^p$ when $2i=s+t$
		\item $\mathcal P^j(xy)=\sum_i\mathcal P^i(x)\mathcal P^{j-i}(y)$
		\item $\beta\mathcal P^j(xy)=\sum_i(\beta\mathcal P^i(x)\mathcal P^{j-i}(y)+\mathcal P^i(x)\beta\mathcal P^{j-i}(y))$
		\item $\mathcal P$ and $\beta\mathcal P$ satisfy the Adem relations.
		\item $\mathcal P^0(\lambda u)=\lambda^p\mathcal P^0(u)$ for all $\lambda \in k$ ($\ch k= p$)
	\end{enumerate}
\end{defn}

In the course of proving things here, the authors adapt these operations to the $\bbZ/2$ grading
by reindexing, calling these new operations $\mathscr P_i$ and $\beta\mathscr P^i$ (they use \texttt{mathscr} but I've
reassigned mine to a font I like better. :P)

Okay, time is running out, so I am pulling together some questions:
\begin{qst}
	Is the cohomology you're using here group cohomology? It seems to me the Lyndon-Hochschild-Serre
	applies to group cohomology.

	If so, I will need to focus some time on getting familiar with it.
\end{qst}

Yes it is group cohomology, or more specifically group \textit{scheme} cohomology. Here we 
are computing $\Ext_G^{i,\varepsilon}(k,k)$.

We begin with our group algebra (here $\Lambda(v_1,v_2)\#k\bbG_{a(1)}$) and take 
a projective resolutions and computing $H^*(\Hom_A(p^\bullet,k))$.

Recall the syzygys, $\Omega^i k$ that occur in a projective resolution of $k$. Then we can compute the cohomology
via the stable module category:
\[\underline{\Hom}_G(\Omega^ik,k)=\Ext_G^{i,\varepsilon}(k,k)\]

There is a result by Jeremy Rickard (rep theory) and by Buchweitz (more generally) that proves

\begin{qst}
	Where is $H^{*,*}(\bbG_a^-,k)$ computed? The papers claim it is $S^{*,*}(V^\sharp)\cong k[\zeta,\eta]$
	where $\zeta$ and $\eta$ are the dual generators corresponding to the generators $u$ and $v$ of $V$,
	the augmentation ideal of $k(\bbG_a^-\times \bbG_a^-)$.
\end{qst}

Again, this is an example of Koszul duality -- use that $k\bbG_a^-=\Lambda(v)$ where $v$ is primitive in degree 1.
But then $k(\bbG_a^-\times\bbG_a^-)\cong\Lambda(v_1,v_2)$. Then applying Koszul duality you get the 
symmetric algebra. Also you need the K\"unneth formula, where as algebras
\[H^*(A\otimes B,k)\cong H^*(A,k)\otimes H^*(A,k).\]

Evans has a book on The groups of cohomology. Here there are computations and resolutions of (cyclic)
groups. There is also a discussion of spectral sequences which is relatively terse (although it defers
somewhat to other sources).

\begin{qst}
	What is the importance of suspensions and what role do they play in this theory? They 
	have come up at several points and I am realizing they are fairly pervasive.

	In particular, I understand the topological definition, but I don't see why they correspond
	to what we want. I am looking at the formula from \cite{detecting-nilpotence}:
	\[H^{1,1}(G,k)=\Hom_{sGr/k}(G^0,\mathbb{G}_a^-)^\pi\]
	where $\pi$ is the group of connected components of $G$.
\end{qst}

\begin{qst}
	In $G=(\bbG_a^-\times\bbG_a^-)\rtimes G_{a(1)}$, where the action is said to be ``nontrivial'',
	why does this seem to imply that the action on one of the copies of $\bbG_a^-$ is trivial?

	The fact that there is a LHS spectral sequence 
	\[H^{*,*}(\bbG_a^-\times \bbG_{a(1)},H^{*,*}(\bbG_a^-,k))\Rightarrow H^{*,*}(G,k)\]
	seems to imply that there is a short exact sequence 
	\[1\to \bbG_a^-\to G\to \bbG_a^-\times \bbG_{a(1)}\to 1\]
	but since the product is direct on the right, doesn't this imply that the action of $\bbG_{a(1)}$ on 
	$\bbG_a^-$ is trivial?
\end{qst}

Path 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{To be filed:}
